---
title: "About"
author: "Jessica Edwards"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(janitor)
```

Schools to compare:

Start with Ivy Leagues first, then hope to add more schools as I continue the project.

Brown University*
```{r brown, echo=FALSE}

brown_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8107/screen/21?school_name=Brown+University") %>% html_nodes("table")

brown_race_ugrad_degrees_2018 <- brown_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

brown_race_ugrad_degrees_2018
```

Columbia University*
```{r columbia, echo=FALSE}

columbia_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8224/screen/21?school_name=Columbia+University") %>% html_nodes("table")

columbia_race_ugrad_degrees_2018 <- columbia_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

columbia_race_ugrad_degrees_2018
```

Cornell University*
```{r cornell, echo=FALSE}

cornell_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8046/screen/21?school_name=Cornell+University") %>% html_nodes("table")

cornell_race_ugrad_degrees_2018 <- cornell_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

cornell_race_ugrad_degrees_2018
```

Dartmouth College*
```{r dartmouth, echo=FALSE}

dartmouth_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8085/screen/21?school_name=Dartmouth+College") %>% html_nodes("table")

dartmouth_race_ugrad_degrees_2018 <- dartmouth_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

dartmouth_race_ugrad_degrees_2018
```

Harvard University*
```{r harvard, echo=FALSE}

harvard_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8130/screen/21?school_name=Harvard+University") %>% html_nodes("table")

harvard_race_ugrad_degrees_2018 <- harvard_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

harvard_race_ugrad_degrees_2018
```

University of Pennsylvania*
```{r penn, echo=FALSE}

penn_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8116/screen/21?school_name=University+of+Pennsylvania") %>% html_nodes("table")

penn_race_ugrad_degrees_2018 <- penn_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

penn_race_ugrad_degrees_2018
```

Princeton University*
```{r princeton, echo=FALSE}

princeton_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8068/screen/21?school_name=Princeton+University") %>% html_nodes("table")

princeton_race_ugrad_degrees_2018 <- princeton_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

princeton_race_ugrad_degrees_2018
```

Yale School of Engineering & Applied Science*
```{r yale, echo=FALSE}

yale_race_ugrad_degrees_2018 <-
  read_html("http://profiles.asee.org/profiles/8218/screen/21?school_name=Yale+School+of+Engineering+%26+Applied+Science") %>% html_nodes("table")

yale_race_ugrad_degrees_2018 <- yale_race_ugrad_degrees_2018[[2]] %>%
  html_table(fill = TRUE)

yale_race_ugrad_degrees_2018
```


Other schools I'd like to consider:
Boston University
California Institute of Technology
University of California, Berkeley
Carnegie Mellon University
Duke University
Georgia Institute of Technology
Harvey Mudd College
Johns Hopkins University
Massachusetts Institute of Technology
Northwestern University
Tufts University
Vanderbilt University
University of Virginia
Washington University in St. Louis